---
title: "BDB_2023_Elo"
author: "Maximilian J. Gebauer"
date: "2025-06-16"
output: html_document
---

```{r Setup, include=FALSE, results='hide', warning=FALSE}
knitr::opts_chunk$set(echo = T, fig.width=8, fig.height=4)
options(scipen = 0, digits = 3)  # controls base R output

# Package setup
if(!require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, dplyr, ggthemes, data.table, lubridate, glmnet,
               GGally, RColorBrewer, ggsci, plotROC, usmap,
               plotly, ggpubr, vistime, coefplot, skimr, car, ggrepel, slider, lubridate,
               tidymodels,ranger,vip,ggplot2, tune,dials,pdp, purrr, stringr, lmtest,
               sandwich)
```

```{r}
#loading in joined data
full_data <- data.table::fread("../../nfl_data/big_data_bowl_2023_AllWeeks.csv")
```

```{r}
#set up response
full_data <- full_data %>%
  mutate(response = ifelse(pff_beatenByDefender == 0 & pff_hitAllowed == 0 & pff_hurryAllowed == 0 & pff_sackAllowed == 0, 1, 0))

#enrusing still have full number of plays from og data post cutting
full_data %>%
  summarise(total_plays=n_distinct(interaction(gameId,playId)))
```

```{r}
#full_test_data <- read.csv("/Users/maximiliangebauer/Downloads/full_test_data.csv")
```

```{r}
#initialize all offense and defense elo ratings with implied defender win odds of 10:1, scaling parameter as 400 (chess)
# assign ratings based on side of ball, ignoring player types unlikely to be in pass blocking interactions
players2 <- read.csv("../../nfl_data/players.csv")
init_play_ratings <- players2 %>% 
  select(nflId, displayName, officialPosition) %>%
  mutate(Defense = as.factor(ifelse(officialPosition %in% c("DE", "DT", "OLB", "ILB", "MLB", "NT", "LB", "SS", "FS", "CB", "DB"), 1, 0)),
         Offense = as.factor(ifelse(officialPosition %in% c("C", "G", "T", "TE", "FB", "WR", "RB", "QB"), 1, 0)),
         Player_Elo = as.numeric(ifelse(Defense=="1",800,1200))
         )
```

```{r}
qb_ids <- players2 %>%
  filter(officialPosition=="QB") %>%
  select(nflId)

qb_ids <- qb_ids$nflId

test_data_filtered <- full_data %>%
  group_by(gameId, playId) %>%
  filter(
    !is.na(pff_nflIdBlockedPlayer) |
    (nflId %in% pff_nflIdBlockedPlayer | nflId %in% qb_ids)
  ) %>%
  ungroup()
```

```{r}
test_data_filtered <- test_data_filtered %>% select(-c(V1,playDirection))
head(test_data_filtered)
```

```{r}
test_data_filtered_snap <- test_data_filtered %>%
  group_by(gameId,playId) %>%
  filter(any(event %in% c("ball_snap"))) %>%
  ungroup()


nrow(test_data_filtered_snap)
nrow(test_data_filtered)
```

```{r}
test_data_filtered_snap %>%
  summarise(total_plays = n_distinct(interaction(gameId, playId)))

test_data_filtered %>%
  summarise(total_plays = n_distinct(interaction(gameId, playId)))
```

```{r}
snap_end_lookup <- test_data_filtered_snap %>%
  group_by(gameId, playId) %>%
  summarise(
    snap_frame = min(frameId[event == "ball_snap"], na.rm = TRUE),
    end_frame  = pmin(
      if (any(event == "pass_forward"))  min(frameId[event == "pass_forward"])  else Inf,
      if (any(event == "qb_strip_sack")) min(frameId[event == "qb_strip_sack"]) else Inf,
      if (any(event == "lateral"))       min(frameId[event == "lateral"])       else Inf,
      if (any(event == "qb_sack"))       min(frameId[event == "qb_sack"])       else Inf,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  # if none of the end events ever occur, just treat end_frame = snap_frame
  mutate(end_frame = ifelse(is.infinite(end_frame), snap_frame, end_frame))


# 2) Now rebuild test_data_post_snap to only include frames between snap and end
test_data_post_snap <- test_data_filtered_snap %>%
  inner_join(snap_end_lookup, by = c("gameId","playId")) %>%
  filter(frameId >= snap_frame,
         frameId <= end_frame) %>%
  ungroup() %>%
  select(-snap_frame, -end_frame)
```

```{r}
head(test_data_post_snap,1000)

test_data_post_snap %>%
  summarise(total_plays = n_distinct(interaction(gameId, playId)))
```



```{r}
test_data_post_snap <- test_data_post_snap %>%
  group_by(gameId,playId) %>%
  filter(any(event %in%c("pass_forward","qb_strip_sack","lateral", "qb_sack"))) %>%
  ungroup()

df <- test_data_post_snap

qb_pos <- df %>%
  filter(officialPosition == "QB") %>%
  transmute(gameId, playId, frameId, x_QB = x, y_QB = y) %>%
  distinct(gameId, playId, frameId, .keep_all = TRUE)

blockers <- df %>%
  filter(!is.na(pff_nflIdBlockedPlayer))%>%
  transmute(gameId, playId, frameId,blocker_id = nflId, rusher_id = pff_nflIdBlockedPlayer) %>% 
  distinct(gameId, playId, frameId, rusher_id, .keep_all = TRUE)

rusher_pos <- df %>%
  semi_join(blockers, by = c("gameId","playId","nflId" = "rusher_id")) %>%
  transmute(gameId, playId, frameId, rusher_id = nflId, x_rusher = x, y_rusher = y) %>%
  distinct(gameId, playId, frameId, rusher_id, .keep_all = TRUE)

blocker_pos <- df %>%
  transmute(gameId, playId, frameId, blocker_id = nflId, x_blocker = x, y_blocker = y) %>%
  distinct(gameId, playId, frameId, blocker_id, .keep_all = TRUE)

# 2) Build pair_metrics (same as before)ŌĆ”
pair_metrics <- blockers %>%
  left_join(rusher_pos,     by = c("gameId","playId","frameId","rusher_id")) %>%
  left_join(qb_pos,         by = c("gameId","playId","frameId")) %>%
  left_join(blocker_pos,    by = c("gameId","playId","frameId","blocker_id")) %>%
  left_join(snap_end_lookup, by = c("gameId","playId")) %>%
  # only keep frames from snap ŌåÆ end_of_play
  filter(frameId >= snap_frame+8, frameId <= end_frame) %>%
  arrange(gameId, playId, rusher_id, blocker_id, frameId) %>%
  group_by(gameId, playId, rusher_id, blocker_id) %>%
  mutate(
    d_rusher    = sqrt((x_rusher  - x_QB)^2 + (y_rusher  - y_QB)^2),
    d_blocker   = sqrt((x_blocker - x_QB)^2 + (y_blocker - y_QB)^2),
    d_prev      = lag(d_rusher),
    strain      = -((d_rusher - d_prev) / 0.1) / d_rusher,
    closer_flag = as.integer(d_rusher < d_blocker),
    ever_closer = as.integer(any(closer_flag == 1, na.rm = TRUE))
  ) %>%
  ungroup()

# 3) A helper for safeŌĆÉmax
safe_max <- function(x) {
  if (all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
}

# 4) Summarise per (game,play,rusher,blocker) without warnings
final_metrics_mod <- pair_metrics %>%
  group_by(gameId, playId, rusher_id, blocker_id) %>%
  summarise(
    mean_strain       = mean(strain, na.rm = TRUE),
    max_strain        = safe_max(strain),
    sd_strain         = sd(strain,   na.rm = TRUE),
    max_zscore_strain = safe_max((strain - mean(strain, na.rm=TRUE)) /
                                   sd(strain, na.rm=TRUE)),
    beat_frame = ifelse(
      any(closer_flag == 1, na.rm = TRUE),
      min(frameId[closer_flag == 1], na.rm = TRUE),
      0
    ),
    rusher_beat_dist = ifelse(
      beat_frame > 0,
      { idx <- which(frameId == beat_frame); d_rusher[idx[1]] },
      NA_real_
    ),
    .groups = "drop"
  )


# 7) Compute each playŌĆÖs ŌĆ£snap_frameŌĆØ as the minimum frameId in test_data_post_snap
snap_lookup <- df %>%
  group_by(gameId, playId) %>%
  summarise(snap_frame = min(frameId), .groups="drop")

# 8) Join & compute time from snap to beat (in seconds)
play_results <- final_metrics_mod %>%
  left_join(snap_lookup, by = c("gameId","playId")) %>%
  mutate(
    time_from_snap_to_beat = pmax((beat_frame - snap_frame) / 10, 0)
  )
```

```{r}
event_frames <- test_data_post_snap %>%
  group_by(gameId, playId) %>%
  summarise(
    pass_forward_frame   = ifelse(any(event == "pass_forward"),   min(frameId[event == "pass_forward"]),   0),
    qb_strip_sack_frame  = ifelse(any(event == "qb_strip_sack"),  min(frameId[event == "qb_strip_sack"]),  0),
    lateral_frame        = ifelse(any(event == "lateral"),        min(frameId[event == "lateral"]),        0),
    qb_sack_frame        = ifelse(any(event == "qb_sack"),        min(frameId[event == "qb_sack"]),        0),
    .groups = "drop"
  )

helper_data <- test_data_post_snap %>%
  rename(blocker_id = nflId) %>%
  select(
    gameId,
    playId,
    blocker_id,
    pff_beatenByDefender,
    pff_hitAllowed,
    pff_hurryAllowed,
    pff_sackAllowed
  ) %>%
  left_join(event_frames, by = c("gameId","playId")) %>%
  distinct()

full_results <- play_results %>%
  left_join(helper_data, by = c("gameId","playId","blocker_id")) %>%
  unique()
```

```{r}
full_results <- full_results %>%
  group_by(gameId, playId, rusher_id) %>%
  mutate(
    n_blockers = n_distinct(blocker_id)
  ) %>%
  ungroup()
```

```{r}
head(full_results,100)
```

```{r}
sub_results <- full_results %>%
  filter(n_blockers==1) %>%
  mutate(
    thresh = factor(
      case_when(
        time_from_snap_to_beat > 0 & time_from_snap_to_beat < 2.5 ~ "1",
        TRUE                                                 ~ "0"
      ),
      levels = c("0","1")
    )
  )

pos_lookup <- test_data_post_snap %>%
  select(gameId, playId, nflId, officialPosition) %>%
  rename(blocker_id = nflId) %>%
  distinct()

sub_results <- sub_results %>%
  left_join(
    pos_lookup,
    by = c("gameId", "playId", "blocker_id")
  )

pos_lookup2 <- test_data_post_snap %>%
  select(gameId, playId, nflId, officialPosition) %>%
  rename(rusher_id = nflId,
         position_rusher = officialPosition) %>%
  distinct()

sub_results <- sub_results %>%
  left_join(
    pos_lookup2,
    by = c("gameId", "playId", "rusher_id")
  )
```

```{r}
sub_results2 %>%
  group_by(as.factor(pff_beatenByDefender)) %>%
  summarise(
    mean_bd = mean(rusher_beat_dist,na.rm=T)
  ) %>%
  ungroup()
```

```{r}
sub_results2<-sub_results%>%filter(pff_hitAllowed==0 & pff_hurryAllowed==0 & pff_sackAllowed==0 & !position_rusher %in% c("CB","G","RB","LB","SS","FS") & !officialPosition %in% c("QB","RB","FB","WR")) %>%
  mutate(beat_dist = ifelse(!is.na(rusher_beat_dist)&rusher_beat_dist<6,1,0)) %>%
  filter(!is.na(beat_dist) & !is.na(sd_strain) & !is.na(max_zscore_strain))
skim(sub_results2)
```
```{r}
write.csv(sub_results2,file="clean_beat_data.csv")
```


```{r}
library(lme4)

model_11 <- glmer(
  pff_beatenByDefender ~ 
    max_zscore_strain + 
    mean_strain + 
    max_strain + 
    sd_strain + 
    officialPosition * thresh +           # both should already be factors
    factor(beat_dist) +                    # if you really want beat_dist as factor
    (1 | position_rusher),                 # random intercept for each rusherŌĆÉposition
  data    = sub_results2,
  family  = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa")
)
```
```{r}
summary(model_11)
```

```{r}
library(caret)

# 1) get probs
probs <- predict(model_11, newdata = sub_results2, type = "response")

# 2) apply threshold
preds <- factor(ifelse(probs > 0.065, 1, 0),
                levels = c(0,1))

# 3) actual labels as factor
actuals <- factor(sub_results2$pff_beatenByDefender,
                  levels = c(0,1))

# 4) simple contingency table
print(table(Predicted = preds, Actual = actuals))

# 5) caret confusion matrix (gives accuracy, sensitivity, etc.)
confusionMatrix(preds, actuals)
```


```{r}
ggplot(full_results,aes(x=as.factor(pff_beatenByDefender),y=max_zscore_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_beatenByDefender),y=max_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_beatenByDefender),y=mean_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_beatenByDefender),y=sd_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_beatenByDefender),y=time_from_snap_to_beat)) +
  geom_boxplot()
```

```{r}
ggplot(full_results,aes(x=as.factor(pff_hitAllowed),y=max_zscore_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hitAllowed),y=max_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hitAllowed),y=mean_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hitAllowed),y=sd_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hitAllowed),y=beat_frame)) +
  geom_boxplot()
```


```{r}
ggplot(full_results,aes(x=as.factor(pff_hurryAllowed),y=max_zscore_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hurryAllowed),y=max_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hurryAllowed),y=mean_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hurryAllowed),y=sd_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_hurryAllowed),y=beat_frame)) +
  geom_boxplot()
```

```{r}
ggplot(full_results,aes(x=as.factor(pff_sackAllowed),y=max_zscore_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_sackAllowed),y=max_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_sackAllowed),y=mean_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_sackAllowed),y=sd_strain)) +
  geom_boxplot()

ggplot(full_results,aes(x=as.factor(pff_sackAllowed),y=beat_frame)) +
  geom_boxplot()
```

```{r}
filter_results <- full_results %>%
  filter(!is.na(pff_beatenByDefender) & !is.na(max_zscore_strain) & !is.na(mean_strain) & !is.na(max_strain) & !is.na(time_from_snap_to_beat)) %>%
  mutate(
    beat_25 = as.factor(ifelse(time_from_snap_to_beat>0&time_from_snap_to_beat <= 2.5,1,0))
  )

pos_lookup <- test_data_post_snap %>%
  select(gameId, playId, nflId, officialPosition) %>%
  rename(blocker_id = nflId) %>%
  distinct()

# then join it in
filter_results <- filter_results %>%
  left_join(
    pos_lookup,
    by = c("gameId", "playId", "blocker_id")
  )

pos_lookup2 <- test_data_post_snap %>%
  select(gameId, playId, nflId, officialPosition) %>%
  rename(rusher_id = nflId,
         position_rusher = officialPosition) %>%
  distinct()

filter_results <- filter_results %>%
  left_join(
    pos_lookup2,
    by = c("gameId", "playId", "rusher_id")
  )
```

```{r}
aaa <- filter_results %>%
  mutate(
    one_five = ifelse(max_zscore_strain > 2,1,0),
    pred = ifelse(one_five==1 | beat_25=="1",1,0)
  )

table(aaa$beat_25,as.factor(aaa$pff_beatenByDefender))
table(as.factor(aaa$one_five),as.factor(aaa$pff_beatenByDefender))
table(as.factor(aaa$pred),as.factor(aaa$pff_beatenByDefender))

filter_results <- filter_results %>%
  mutate(
    response = ifelse((pff_beatenByDefender==1 | pff_hitAllowed==1),1,0)
  )
```

```{r}
model_1 <- glm(as.factor(response) ~ max_zscore_strain + splines::bs(time_from_snap_to_beat,df=6,degree=3) + as.factor(officialPosition), data=filter_results, family = "binomial")
summary(model_1)
probs <- model_1$fitted.values
preds <- as.factor(ifelse(probs>.05,1,0))
table(preds,as.factor(filter_results$pff_beatenByDefender))
```

```{r}
filter_results1 <- filter_results %>%
  mutate(
    response = case_when(
      pff_beatenByDefender == 1 ~ "beaten",
      pff_sackAllowed        == 1 ~ "sack",
      pff_hitAllowed         == 1 ~ "hit",
      pff_hurryAllowed       == 1 ~ "hurry",
      TRUE                        ~ "won"
    )
  )
set.seed(321)
multi_model1 <- nnet::multinom(as.factor(response)~ max_zscore_strain + splines::bs(time_from_snap_to_beat,df=6,degree=3) + as.factor(officialPosition), data=filter_results1)
```
```{r}
filter_results2 <- filter_results1 %>%
  filter(n_blockers==1)

X <- model.matrix(pff_beatenByDefender ~ max_zscore_strain + time_from_snap_to_beat + officialPosition + position_rusher, data=filter_results2)
Y <- as.factor(filter_results2$pff_beatenByDefender)
set.seed(321)
cv_model <- glmnet::cv.glmnet(X,Y,nfolds=10,family="binomial")
```

```{r}
preds <- predict(cv_model,
                 newx = X,
                 s    = "lambda.min",
                 type = "class")

# 2) Make sure itŌĆÖs a factor with the same levels as Y
preds <- factor(as.vector(preds), levels = levels(Y))

# 3) Build the confusion table
conf_mat <- table(Predicted = preds, Actual = Y)
print(conf_mat)
```

```{r}
filter_results3 <- filter_results2 %>%
  mutate(
    beat_3 = as.factor(ifelse(time_from_snap_to_beat>0&time_from_snap_to_beat <= 3.0,1,0))
  )
model_2 <- glm(as.factor(pff_beatenByDefender) ~  as.factor(beat_3) + as.factor(officialPosition) + as.factor(position_rusher), data=filter_results3, family = "binomial")
summary(model_2)
probs <- model_2$fitted.values
preds <- as.factor(ifelse(probs>.1,1,0))
table(preds,as.factor(filter_results2$pff_beatenByDefender))
```

```{r}
export_data <- filter_results1 %>%
  select(gameId,playId,rusher_id,blocker_id,mean_strain,sd_strain,max_strain,max_zscore_strain,time_from_snap_to_beat,n_blockers,officialPosition,position_rusher,response)
```

```{r}
write.csv(export_data,file = "predict_beat_data.csv")
```

```{r}
test_data_post_snap %>%
  group_by(gameId,playId) %>%
  filter(any(event %in%c("pass_forward","qb_strip_sack","lateral", "qb_sack")))
```

for the code setup below. Within each joint grouping by  (gameId, playId, and nflId), there are a number of observations by frameId (numeric, taking integer values). There is an "event" column in the data frame that takes various character values or NA if no event occurred in that frame.  
```{r}
post_snap <- test_data_filtered %>%
  group_by(gameId,playId,nflId) 

```



```{r}
interactions <- test_data_filtered %>%
  filter(!is.na(pff_nflIdBlockedPlayer)) %>%
  select(gameId, playId,
         blocker_nflId = nflId,
         rusher_nflId  = pff_nflIdBlockedPlayer,
         outcome = response,
         everything())

pass_rusher_positions <- c("DE", "DT", "OLB", "ILB", "NT", "SS", "DB", "LB", "MLB", "FS", "CB")

valid_rushers <- init_play_ratings %>%
  filter(officialPosition %in% pass_rusher_positions) %>%
  pull(nflId) %>% 
  as.character()

interactions_pr <- interactions %>%
  filter(rusher_nflId %in% valid_rushers)
```

```{r}
init_ratings <- init_play_ratings %>%
  select(nflId, Player_Elo)

elo_vec <- setNames(init_ratings$Player_Elo,
                    as.character(init_ratings$nflId))

learning_rate     <- 32

scaling_parameter <- 400
```

```{r}
unique(interactions_pr$gameId)
selected_games <- c("2021090900", "2021091200", "2021091201", "2021091202", "2021091203", "2021091204", "2021091205", "2021091206", "2021091207", "2021091208", "2021091209")

elo_update_function <- function(blocker_elo, rusher_elo, learning_rate, scaling_parameter, outcome) {
  expected_blocker_win <- 1 / (1 + 10^((rusher_elo - blocker_elo) / scaling_parameter))
  expected_rusher_win <- 1 - expected_blocker_win

  new_blocker_elo <- blocker_elo + learning_rate * (outcome - expected_blocker_win)
  new_rusher_elo  <- rusher_elo  + learning_rate * ((1 - outcome) - expected_rusher_win)

  list(
    blocker_elo = new_blocker_elo,
    rusher_elo  = new_rusher_elo
  )
}

filtered_games <- interactions_pr %>%
  filter(as.character(gameId) %in% as.character(selected_games))

history <- filtered_games %>%
  mutate(
    before_blocker_elo = NA_real_,
    before_rusher_elo  = NA_real_,
    after_blocker_elo  = NA_real_,
    after_rusher_elo   = NA_real_
  )

for (i in seq_len(nrow(history))) {
  row    <- history[i, ]
  b_id   <- as.character(row$blocker_nflId)
  r_id   <- as.character(row$rusher_nflId)
  out    <- row$outcome 

  b_elo  <- elo_vec[b_id]
  r_elo  <- elo_vec[r_id]

  new_elos <- elo_update_function(
    blocker_elo       = b_elo,
    rusher_elo        = r_elo,
    learning_rate     = learning_rate,
    scaling_parameter = scaling_parameter,
    outcome           = out
  )

  history$before_blocker_elo[i] <- b_elo
  history$before_rusher_elo[i]  <- r_elo
  elo_vec[b_id] <- new_elos$blocker_elo
  elo_vec[r_id] <- new_elos$rusher_elo
  history$after_blocker_elo[i]  <- new_elos$blocker_elo
  history$after_rusher_elo[i]   <- new_elos$rusher_elo
}

```

```{r}
str(interactions_pr$gameId)
unique(as.character(interactions_pr$gameId))
intersect(as.character(interactions_pr$gameId), as.character(selected_games))
```

```{r}
history2 <- history %>%
  mutate(interaction = row_number())

elo_tracks <- bind_rows(
  history2 %>%
    transmute(interaction,
              nflId   = as.character(blocker_nflId),
              elo     = after_blocker_elo,
              role    = "blocker"),
  history2 %>%
    transmute(interaction,
              nflId   = as.character(rusher_nflId),
              elo     = after_rusher_elo,
              role    = "rusher")
)

ggplot(elo_tracks, aes(x = interaction, y = elo, group = nflId, color = role)) +
  geom_line(alpha = 0.7) +
  labs(x = "Interaction #",
       y = "Elo rating",
       color = "Role") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
final_elos <- tibble(
  nflId     = names(elo_vec),
  final_elo = as.numeric(elo_vec)
)

players_elos <- init_play_ratings %>%
  select(nflId, displayName, officialPosition, Defense, Offense) %>%
  mutate(nflId = as.character(nflId)) %>%
  left_join(final_elos, by = "nflId")

rusher_ids  <- unique(interactions_pr$rusher_nflId)
blocker_ids <- unique(interactions_pr$blocker_nflId)

top_rushers <- players_elos %>%
  filter(nflId %in% rusher_ids) %>%
  arrange(desc(final_elo)) %>%
  slice_head(n = 10)

top_blockers <- players_elos %>%
  filter(nflId %in% blocker_ids) %>%
  arrange(desc(final_elo)) %>%
  slice_head(n = 10)

top_rushers
top_blockers
```

```{r}
history2 <- history %>%
  mutate(interaction = row_number())

# 2) Stack blocker + rusher tracks into one long table
elo_tracks <- bind_rows(
  history2 %>%
    transmute(
      interaction,
      nflId = as.character(blocker_nflId),
      elo   = after_blocker_elo,
      role  = "blocker"
    ),
  history2 %>%
    transmute(
      interaction,
      nflId = as.character(rusher_nflId),
      elo   = after_rusher_elo,
      role  = "rusher"
    )
)
tracks_named <- elo_tracks %>%
  left_join(
    init_play_ratings %>%
      mutate(nflId = as.character(nflId)) %>%
      select(nflId, displayName),
    by = c("nflId")
  )

# 2) Top-10 rushers
top_rusher_ids <- top_rushers %>% pull(nflId) %>% as.character()

rusher_tracks <- tracks_named %>%
  filter(role == "rusher", nflId %in% top_rusher_ids)

ggplot(rusher_tracks, aes(x = interaction, y = elo, color = displayName)) +
  geom_line(size = 1) +
  labs(
    title = "Elo Trajectories of Top 10 Pass-Rushers",
    x     = "Interaction #",
    y     = "Elo Rating",
    color = "Player"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# 3) Top-10 blockers
top_blocker_ids <- top_blockers %>% pull(nflId) %>% as.character()

blocker_tracks <- tracks_named %>%
  filter(role == "blocker", nflId %in% top_blocker_ids)

ggplot(blocker_tracks, aes(x = interaction, y = elo, color = displayName)) +
  geom_line(size = 1) +
  labs(
    title = "Elo Trajectories of Top 10 Blockers",
    x     = "Interaction #",
    y     = "Elo Rating",
    color = "Player"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```


```{r}
top_blockers_pos <- players_elos %>%
  filter(nflId %in% blocker_ids & !(officialPosition %in% c("QB", "WR", "RB"))) %>%
  group_by(officialPosition) %>%
  arrange(desc(final_elo)) %>%
  slice_head(n = 10)

top_blockers_pos

top_rushers_post <- players_elos %>%
  filter(nflId %in% rusher_ids) %>%
  group_by(officialPosition) %>%
  arrange(desc(final_elo)) %>%
  slice_head(n = 10) %>%
  filter(!officialPosition %in% c("CB", "SS", "FS"))

top_rushers_post
```

```{r}
new_data <- fread("/Users/maximiliangebauer/Downloads/Hudl IQ data for Wharton School/Hudl IQ 2021 NFL Events.csv")
```

```{r}
names(new_data)
head(new_data)
```
snap, engagement start, engagement end, 
```{r}
new_data %>%
  filter(event_types %in% c("{\"\"Engagement Start\"\"}","{\"\"Engagement End\"\"}") & event_player_name=="Dexter Lawrence" & opponent_player_name=="Matt Hennessy")
```
sub_gameid <- game_id %>%
  group_by(gameId,playId) %>%
  filter(any(event %in% c("pass_forward", "pass_shovel", "pass_tipped", "qb_sack",
                          "pass_outcome_touchdown", "pass_arrived", "qb_kneel",
                          "pass_outcome_caught", "pass_outcome_incomplete"))) %>%
  ungroup()
```{r}
sample_game <- new_data %>%
  filter(game_id == 1069714)

length(unique(new_data$play_uuid))

pass_data <- new_data %>% 
  group_by(play_uuid) %>%
  filter(any(event_types %in% c("{Pass}","{Passer-Hit}","{Pass,Hit}","{Pass,Off-Camera}","{Tackled,Sacked}","{Sacked,Tackled}")))

length(unique(pass_data$play_uuid))
```

1024025 Dexter Lawrence
```{r}
sample_game %>%
  filter(play_uuid=="a0dc69e6-c5f9-4a81-a823-e92a3efc918c")
```

```{r}
head(new_data)
```

```{r}

```






